---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EMODnetWCS

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/EMODnetWCS)](https://CRAN.R-project.org/package=EMODnetWCS)
<!-- badges: end -->


The goal of EMODnetWCS is to allow interrogation of and access to EMODnet geographic raster data in R though the [EMODnet Web Coverage Services](https://www.emodnet.eu/en/data). [Web Coverage services (WCS)](https://www.ogc.org/standards/wcs) offer multi-dimensional coverage data for access over the Internet. This package was developed by Sheffield University as part of EMODnet Biology WP4.



```{r, eval=FALSE}
vignette("emodnetwfs")
```


## Installation

You can install the development version of EMODnetWCS from GitHub with:

``` r
# install.packages("remotes")
remotes::install_github("EMODnet/EMODnetWCS")
```

## Available services

All available services are contained in the tibble returned by `emodnet_wcs()`.


```{r, echo=FALSE}
library(EMODnetWCS)
knitr::kable(emodnet_wcs())
```

To explore available services in Rstudio use:

```r
View(emodnet_wcs())
```


## Create Service Client

Create new WCS Client.  Specify the service using the `service` argument. 

```{r}
wcs_bio <- emodnet_init_wcs_client(service = "biology")

wcs_bio$getUrl()
```

