---
title: "Downloading Coverages"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

vcr::setup_knitr()
```

emodnet.wcs provides functionalitly for downloading coverage data from EMODnet WCS services as `terra` [`<spatRaster>`](https://rspatial.github.io/terra/reference/SpatRaster-class.html) class objects. It also allows subsetting along coverage dimensions as well as requesting specific bands for download.

For more details on extracting metadata to inform valid subsetting of coverages, please refer to the [Getting metadata about Services & Coverages](https://emodnet.github.io/emodnet.wcs/articles/metadata.html) article in the `emodnet.wcs` online documentation.


The first step to downloading coverages is to initiate a connection with on of the EMODnet WCS servers.

```{r setup}
library(emodnet.wcs)
```


```{r}
#| cassette: bio
wcs <- emdn_init_wcs_client(service = "biology")

```

You can get a list of available coverages from the service with `emdn_get_coverage_ids()`

```{r}
#| cassette: coverage-ids
coverage_ids <- emdn_get_coverage_ids(wcs)
coverage_ids
```

Let's focus on the _Calanus finmarchicus_
```{r}
#| cassette: cov3
coverage_id <- coverage_ids[3]
coverage_id

```


## Subset using a bounding box

While downloading full coverages for most services will work, it may take a long time depending on the size of the coverage so it's often a good idea to focus on areas of interest and use bounding boxes to subset requests.

<div class="alert alert-warning" role="alert">

<h5> **Warning!**</h5>

<p>While most coverages can be downloaded in full, the coverages from the `"bathymetry"` service in particular are 100GB in size and full download will not be permitted by the server so spatial subsetting **must** be used with the services. Otherwise the request will return a `curl` error (most likely ` curl: (92) HTTP/2 stream 0 was not closed cleanly: INTERNAL_ERROR (err 2)`). Note that, in the near future, it will be possible to download larger boxes than what the WCS can currently handle, but this will still be capped at 2 GB.</p>

<p>If you need to download the full [EMODnet Digital Bathymetry (DTM)](https://data.europa.eu/data/datasets/emodnet_bathymetry?locale=en), you will have to download it tile by tile from <https://portal.emodnet-bathymetry.eu/></p>

</div>

So let's also create a subsetting `bbox` of an area of interest, the southern North Sea. 

We can first check the bounding box of the full coverage to ensure the subset bounding box lies within.

```{r}
#| cassette: summary
cov_summary <- emdn_get_coverage_summaries(
  wcs,
  coverage_id
)[[1]]

emdn_get_bbox(cov_summary)
```

We can then create a bounding ensuring it lies within the coverage's bounding box.
```{r}
bbox <- c(xmin = 1.5, ymin = 53, xmax = 6.5, ymax = 58)
```


Now we are ready to download the coverage and provide the bounding box we created to the `bbox` argument.

```{r}
#| cassette: get-cov1
cov_1 <- emdn_get_coverage(wcs, coverage_id = coverage_id, bbox = bbox)

cov_1
```

```{r}
terra::plot(cov_1)
```

The raster returned from the server includes the nil value (which technically encodes `NA`) as a number.

You can access nil values for each band from a `summary` object using `emdn_get_band_nil_values()`

```{r}
#| cassette: get-nil
emdn_get_band_nil_values(cov_summary)
```

If you wish to encode nil values as `NA` during coverage download, you can use argument `nil_values_as_na`.

```{r}
#| cassette: get-cov-nil
cov_nil_to_na <- emdn_get_coverage(
  wcs,
  coverage_id = coverage_id,
  bbox = bbox,
  nil_values_as_na = TRUE
)

cov_nil_to_na
```

```{r}
terra::plot(cov_nil_to_na)
```

Now, nil values are encode as `NA` and plotting is much more informative as the range of values within the raster is now constrained to valid values of interest.

## Subset using a bounding box and specific timepoints

If a coverage has additional dimensions to the two geographic ones (x/long & y/lat) then the last point on that dimension is always returned if none are specified through the request.

Coverage `r coverage_id` indeed has a temporal extent:

```{r}
#| cassette: dim-types
emdn_get_dimension_types(cov_summary)
```

so our previous request returned only the last timepoint of the coverage:

However we can request specific timepoints. First, let's get a list of all available timepoints.

```{r}
#| cassette: get-temp
temp_coefs <- emdn_get_coverage_dim_coefs(
  wcs = wcs,
  coverage_ids = coverage_id,
  type = "temporal"
)
```

Next, let's select the first 3 available timepoints.
```{r}
timepoints <- head(temp_coefs[[1]], 3)
timepoints
```

We can request the specific timepoints by supplying our vector of timepoints to the `time` argument.
```{r}
#| cassette: get-cov2
cov_2 <- emdn_get_coverage(
  wcs,
  coverage_id = coverage_id,
  bbox = bbox,
  time = timepoints,
  nil_values_as_na = TRUE
)
```

```{r}
cov_2
```

Note that this returns a ` <SpatRaster>` Stack instead of a single `<SpatRaster>`.
```{r}
terra::plot(cov_2)
```


A similar approach is used for specifying points in the vertical dimension, using argument `elevation` where appropriate.

## Subset a specific band

We can also request specific bands from a coverage. 

First let's check the available bands in our coverage.

```{r}
#| cassette: get-band-desc
emdn_get_band_descriptions(cov_summary)
```

To download a specifc band we use the argument `rangesubset`:
```{r, eval=FALSE}
#| cassette: get-subset
emdn_get_coverage(
  wcs,
  coverage_id = coverage_id,
  bbox = bbox,
  rangesubset = "Relative error",
  nil_values_as_na = TRUE
) %>%
  terra::plot()
```



## Specify an output filename

When downloading data, underlying package `ows4R` uses the in-built [`WCSCoverageFilenameHandler()`](https://eblondel.github.io/ows4R/reference/WCSCoverageFilenameHandler.html) function for generating appropriate filenames from the parameters supplied to the request.

However, it is possible to override this as well as the location the file is written to by supplying a file path to argument `filename`.

```{r}
#| cassette: tif
emdn_get_coverage(
  wcs,
  coverage_id = coverage_id,
  bbox = bbox,
  rangesubset = "Relative error",
  filename = "cal-fin_rel-error.tif"
)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
fs::dir_ls(type = "file", glob = "*.tif") %>%
  fs::file_delete()
```

